Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint

### 2014-11-13 intro.tex

Recently, an even higher level of geometric detail was reached in the
form of aligning
3D CAD model instances to real world test images~\cite{Aubry14,
  Lim14,Kholgade14, Chen13, Kostas14}. Interestingly, the problem of matching 3D models to 2D
images has been targeted since the early days of computer
vision~\cite{Lowe87}, but had largely been neglected in recent years
in favor of 2D detectors based on robust local features and
statistical learning techniques. Now, this problem is being revisited
mostly due to two reasons: the availability of 3D CAD models for many
object classes and the availability of robust image matching
techniques.
%
% performance object detectors usually generalizes very well. It
% can handle various intraclass variability, viewpoint change and
% occlusion. However due to this high generalization, sometimes it is
% very difficult to get accurate pose, category or 2D-3D matching. Our
% method, specialized in giving such accurate information, can be
% applied on any generic object detectors to give high quality metadata.
% %
% Matching 3D CAD model to an image has been a classical problem that
% dates back to eighties . Lowe \etal , the vision
% community instead focused on the object detection tasks that estimates
% a rectangular bounding box around an object\cite{Felzenszwalb10,
%   Girshick14}. 
%
%In addition, there has been a growing emand for accurate 2D-3D
%matching \cite{Kholgade14, Chen13, Kostas14} for image manipulation,
%editing, and data augmentation in the computer graphics community.
%
% Recently, with the powerful computing capabilities and a large
% collection of CAD models, the 2D-3D matching (registration) problem
% gets more and more approachable but due to inherent difficulty and
% limited performance of existing methods, many applications still
% require supervision to achieve 2D-3D matching.
% In medical community, registering 3D scan data to 2D image enabled various minimally invasive surgeries \cite{Markelj12}.
%
% For category-level matching, \cite{Xiang12, Hejrati14, Pepik12} used
% 3D parts or aspect to represent underlying category-level 3D geometry
% of the object. With the advent of large public CAD datasets
% \cite{Trimble, Yobi}, the instance-level object detection and 2D-3D
% registration became possible. \cite{Aubry14, Lim14} used large
% collection of CAD models to faithfully align 3D CAD model to an image
% by rendering CAD models and creating 2D templates for various
% viewpoints.

Recent approaches to 2D-3D matching typically represent an object
class as a collection of 3D exemplar models, which they render from a
large number of viewpoints. The resulting artifial images are then
used to train exemplar models that can be matched to a real-world
image at test time.

\cite{Aubry14, Lim14} require a large number of rendering and 2D
templates for each of the CAD model with specific viewpoint. Since
there are so many templates that it becomes almost impossible to get
all the true positive image from training data. Instead \cite{Aubry14,
  Lim14} used whitening as a preprocessing stage \cite{Hariharan12} to
make templates more discriminative. Then calibration is done on a
small training dataset.

Although whitening to create template is faster than some of
instance-level matching algorithms such as training Exemplar-SVM
\cite{Malisiewicz11} or complicated models such as LEVAN
\cite{Divvala14}, whitening and calibration large number of templates
takes very long time. Also, once the fixed discrete viewpoint
templates have been made and calibrated, it is difficult to fine-tune
the prediction made using the large collection of templates.

@@

In our paper, we present Non-Zero Whiten Histograms of Orientations
(NZ-WHO) to dramatically speed up the template generation and get rid
of time consuming calibration and yet achieves performance on par with
calibrated templates. First, we will present NZ-WHO is more
discriminative than WHO and introduce Conjugate Gradient method to
speed up the whitening and create a high resolution template in just
80 milliseconds. Also, using this fast NZ-WHO template generation, we
present \textbf{on-the-fly} template generation to fine tune an
initial pose estimation to continuously refine translation, viewpoints
(yaw, roll, pitch) and focal length.

>>>>

Recently, an even higher level of geometric detail was reached in the
form of aligning
3D CAD model instances to real world test images~\cite{Aubry14,
  Lim14,Kholgade14, Chen13, Kostas14}. Interestingly, the problem of matching 3D models to 2D
images has been targeted since the early days of computer
vision~\cite{Lowe87}, but had largely been neglected in recent years
in favor of 2D detectors based on robust local features and
statistical learning techniques. Now, this problem is being revisited
mostly due to two reasons: {\em (i)} the availability of 3D CAD models
for many object classes and {\em (ii)} the availability of robust
image matching techniques.
%
% performance object detectors usually generalizes very well. It
% can handle various intraclass variability, viewpoint change and
% occlusion. However due to this high generalization, sometimes it is
% very difficult to get accurate pose, category or 2D-3D matching. Our
% method, specialized in giving such accurate information, can be
% applied on any generic object detectors to give high quality metadata.
% %
% Matching 3D CAD model to an image has been a classical problem that
% dates back to eighties . Lowe \etal , the vision
% community instead focused on the object detection tasks that estimates
% a rectangular bounding box around an object\cite{Felzenszwalb10,
%   Girshick14}. 
%
%In addition, there has been a growing emand for accurate 2D-3D
%matching \cite{Kholgade14, Chen13, Kostas14} for image manipulation,
%editing, and data augmentation in the computer graphics community.
%
% Recently, with the powerful computing capabilities and a large
% collection of CAD models, the 2D-3D matching (registration) problem
% gets more and more approachable but due to inherent difficulty and
% limited performance of existing methods, many applications still
% require supervision to achieve 2D-3D matching.
% In medical community, registering 3D scan data to 2D image enabled various minimally invasive surgeries \cite{Markelj12}.
%
% For category-level matching, \cite{Xiang12, Hejrati14, Pepik12} used
% 3D parts or aspect to represent underlying category-level 3D geometry
% of the object. With the advent of large public CAD datasets
% \cite{Trimble, Yobi}, the instance-level object detection and 2D-3D
% registration became possible. \cite{Aubry14, Lim14} used large
% collection of CAD models to faithfully align 3D CAD model to an image
% by rendering CAD models and creating 2D templates for various
% viewpoints.

For (i), recent approaches to 2D-3D matching~\cite{Aubry14, Lim14}
typically rely on a
collection of 3D exemplar models, which they render from a
large number of viewpoints. The resulting artifial images are then
used to train exemplar models that can be matched to a real-world
image at test time.
%
For (ii), it has been realized that template-based exemplar detectors
based on HOG~\cite{Dalal05} features can be trained analytically, by
replacing the standard SVM with an LDA classifier
(E-LDA)~\cite{Hariharan12}. As a result, it is feasible to train
hundreds of thousands of mid-level patch detectors for
recognition~\cite{Aubry14}.
%
Unfortunately, the performance of E-LDA relies crucially on an
additional calibration step that equalizes the detection scores of
independently trained exemplar models. Since this step involves costly
mining of hard negative examples~\cite{Dalal05,Felzenszwalb10} on a
validation set, it constitutes the major computational bottleneck of
E-LDA, limiting its scalability.

In this paper, we propose a novel method for 2D-3D alignment of
exemplar CAD models to real-world images that circumvents the need for
calibration, and greatly enhances the scalability of E-LDA. As a
result, we can render novel views and train corresponding exemplar
models {\em on the fly}, without the need for offline processing. As a
by-product, we can formulate the alignment problem as a parameter
search in a continuos pose space, consisting of azimuth, elevation,
and ~\scream{}, resulting in truely continuous pose estimates, which
we implement using MCMC sampling.
%  require a large number of rendering and 2D
% templates for each of the CAD model with specific viewpoint. Since
% there are so many templates that it becomes almost impossible to get
% all the true positive image from training data. Instead \cite{Aubry14,
%   Lim14} used whitening as a preprocessing stage \cite{Hariharan12} to
% make templates more discriminative. Then calibration is done on a
% small training dataset.
%
% Although whitening to create template is faster than some of
% instance-level matching algorithms such as training Exemplar-SVM
% \cite{Malisiewicz11} or complicated models such as LEVAN
% \cite{Divvala14}, whitening and calibration large number of templates
% takes very long time. Also, once the fixed discrete viewpoint
% templates have been made and calibrated, it is difficult to fine-tune
% the prediction made using the large collection of templates.

Our paper makes the following specific contributions.

First, we present a novel method for training exemplar models from
rendered 3D CAD data {\em on the fly}, enabled by a novel variant of
E-LDA, and making efficient use of the specific characteristics of
rendered images. To our knowledge, our method constitutes the first
attempt to simultanously render and train exemplar detectors on the
fly.
% In our paper, we present Non-Zero Whiten Histograms of Orientations
% (NZ-WHO) to dramatically speed up the template generation and get rid
% of time consuming calibration and yet achieves performance on par with
% calibrated templates. First, we will present NZ-WHO is more
% discriminative than WHO and introduce Conjugate Gradient method to
% speed up the whitening and create a high resolution template in just
% 80 milliseconds. Also, using this fast NZ-WHO template generation, we
% present \textbf{on-the-fly} template generation to fine tune an
% initial pose estimation to continuously refine translation, viewpoints
% (yaw, roll, pitch) and focal length.

Second, we show to enrich the output of an existing object class
detector, such as the DPM~\cite{Felzenszwalb10} or the
R-CNN~\cite{Girshick14} with additional 3D information, applying our
method to candidate detections provided by the respective detector. As
a result, we can refine the original detections with additional
estimates of 3D continuos pose and a 3D CAD model exemplar.

And third, we give an in-depth experimental study that demonstrates
the effectiveness of our approach on a standard benchmark for
object detection and viewpoint estimation~\cite{Xiang14}.

outperforming previous results on this dataset by considerable
margins.

@@
